### 2021-12-26 Paper_online_Learning.Rmd
update introduction

---
title: "Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends"
author:
- familyname: Talagala 
  othernames: Priyanga Dilini
  address: Department of Compuational Mathematics, University of Moratuwa, Sri Lanka
  email: priyangad@uom.lk
  correspondingauthor: true
- familyname: Talagala
  othernames: Thiyanga S.
  address: Department of Statistics, Faculty of Applied Sciences, University of Sri Jayewardenepura, Sri Lanka
  email: ttalagala@sjp.ac.lk
abstract: "Distance Education has a long history. However COVID-19 has made a new era of distance education. With the increasing demand, various different distance learning solutions have been introduced for different education purposes. In this study we invetigated the impact of COVID-19 on ditance education and the possibility of using Google Trend search queries as a proxy to quantify the  popularity and the public interest towards different distance education solutions. Both visual and analytical approaches were used to  analyse global level web search querie during COVID-19 pandemic. The findings of this study include a Google Trend footprint of various distance leraning solutions for diffrent education purposes that can be used as a proxy for measuring global attention towards different distance learning solutions. The can provide a fast first step guide to identify the most popular distance learning tools available for different educational purposes. R code and data to reproduce the results of this work are available in the online supplementary materials."
keywords:  "Online-learning, Distance education solutions, COVID-19 Pandemic,  Google Trend search queries "
wpnumber: 1
jelcodes: C82,I20,I21
blind: false
cover: true
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
output:
  templatehub::workingpaper:
    fig_caption: yes
    fig_height: 10
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
```

```{r main, child='main.Rmd'}
```