### 2021-12-26 Paper_online_Learning.Rmd
edit typos

---
title: "COVID-19 and Online Learning Tools"
author:
- familyname: Talagala 
  othernames: Priyanga Dilini
  address: Department of Compuational Mathematics, University of Moratuwa, Sri Lanka
  email: priyangad@uom.lk
  correspondingauthor: true
- familyname: Talagala
  othernames: Thiyanga S.
  address: Department of Statistics, Faculty of Applied Sciences, University of Sri Jayewardenepura, Sri Lanka
  email: ttalagala@sjp.ac.lk
abstract: "Distance Education has a long history. However COVID-19 has made a new era of distance education. Due to the  increasing demand, various distance learning solutions have been introduced for different distance education purposes. In this study we investigated the impact of COVID-19 on distance education and the possibility of using Google Trend search queries as a proxy to quantify the  popularity and the public interest towards different distance education solutions. Both visual and analytical approaches were used to  analyse global level web search queries during COVID-19 pandemic. The findings of this study include a Google Trend footprint of various distance leraning solutions that can be used as a proxy for measuring global attention towards different distance learning solutions. This can provide a fast first step guide to identify the most popular online learning tools available for different educational purposes. R code and data to reproduce the results of this work are available in the online supplementary materials."
keywords:  "Online learning, Online teaching Distance education solutions, COVID-19 Pandemic,  Google Trend search queries"
wpnumber: 1
jelcodes: C82,I20,I21
blind: false
cover: true
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
output:
  templatehub::workingpaper:
    fig_caption: yes
    fig_height: 10
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
```

```{r main, child='main.Rmd'}
```